Clustering time series based on probability distributions across temporal granularities

### 2021-10-31 main.Rmd
methodology done

- _Data preparation_

@wang2020tsibble introduced the tidy "tsibble" data structure to support exploration and modeling of temporal data comprising of an index, optional key(s), and measured variables. <!-- An index is a variable with inherent ordering from past to present and a key is a set of variables that define observational units over time. A linear granularity is a mapping of the index set to subsets of the time domain. For example, if the index of a tsibble is days, then a linear granularity might be weeks, months or years.--> For each key variable, the raw smart meter data is a sequence that is indexed by time and comprises values of several measurement variables at each time point. This sequence, though, could be depicted in a variety of ways. A shuffling of the raw sequence could reflect the distribution of hourly consumption over a single day, while another could indicate consumption over a week or a year. These temporal deconstructions of a time period into units such as hour-of-day, work-day/weekend are called cyclic temporal granularities. All cyclic granularities can be expressed in terms of the index set and could be augmented with the initial tsibble structure (index, key, measurements). It is worthwhile to note that the data structure changes while transporting from linear to cyclic scale of time as multiple observations of the measured variable would correspond to each category of the cyclic granularities. In this paper, quantiles are chosen to characterize the distributions for each category of the cyclic granularity. So, each category of a cyclic granularity corresponds to a list of numbers which is essentially few chosen quantiles of the multiple observations.

@@

In the analysis of energy smart meter data, K-Means or hierarchical clustering are often employed. These are simple and effective techniques that work well in a range of scenarios. For clustering, both employ a distance measure, and the distance measure chosen has a major influence on the structure of the clusters. We employ agglomerative hierarchical clustering in conjunction with Ward's criteria (XXX reference). The pair of clusters with minimum between-cluster distance are are sequentially merged in this using this agglomerative algorithms. A good comprehensive list of algorithms can be found in @.Xu2015-ja. We can possibly employ any clustering method that supports the given distance metric as input.

>>>>

- _Data pre-preprocessing_

@wang2020tsibble introduced the tidy "tsibble" data structure to assist temporal data exploration and modeling. To start with, the measured variable for each key variable (observation in this context) is a time-indexed sequence of values for various measurement variables at each time point. This sequence, however, could be shown in several ways. A shuffle of the raw sequence may represent hourly consumption throughout a day, a week, or a year. Cyclic granularities like hour-of-day, hour-of-week can be expressed in terms of the index set in the "tsibble" data structure. But the data structure changes while transporting from linear to cyclic scale of time as multiple observations of the measured variable would correspond to each category of the cyclic granularities. In this paper, quantiles are chosen to characterize the distributions induced by the multiple observations for each category of the cyclic granularity. So, each category of a cyclic granularity corresponds to a list of numbers which is essentially few chosen quantiles of the multiple observations.

<!-- An index is a variable with inherent ordering from past to present and a key is a set of variables that define observational units over time. A linear granularity is a mapping of the index set to subsets of the time domain. For example, if the index of a tsibble is days, then a linear granularity might be weeks, months or years.-->

@@

With a way to obtain pairwise distances, any clustering algorithm can be employed that supports the given distance metric as input. A good comprehensive list of algorithms can be found in @Xu2015-ja based on traditional ways like partition, hierarchy or more recent approaches like distribution, density and others. We employ agglomerative hierarchical clustering in conjunction with Ward's criteria (XXX reference). Hierarchical cluster techniques fuse neighbouring points sequentially to form bigger clusters, beginning with a full pairwise distance matrix. The distance between clusters is described using a "linkage technique". This agglomerative approach successively merges the pair of clusters with the shortest between-cluster distance using Ward's linkage method. Hierarchical algorithms is one of the most widely used, can operate with data of any form, has reasonable scalability, and the number of clusters is not needed as a parameter.
